---
title: "Habitat conservation goal based on MAR"
output: html_fragment
params:
  mode: NULL
  structure: NULL
  aloss: NULL
  gtarg: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
```

# Overview

This report summarizes predictions based on the **mutations–area relationship (MAR)**, which describes how genetic diversity scales with geographic area. The MAR follows a power-law:

$$
M = cA^z
$$

where:

- $M$ is the genetic diversity measured as the number of segregating mutations,

- $A$ is the area sampled for genetic data,

- $z$ is the scaling exponent, influenced by population structure, demographic history, and species biological traits.

This can be rearranged to estimate habitat conservation goals given genetic diversity goals:

$$
\text{% Genetic diversity to preserve} = (1 - \text{% Maximum habitat loss})^z
$$

## User Input

- **Mode**: `r params$mode`
- **Population structure (z)**: `r params$structure`
- **Percentage of genetic diversity to preserve**: `r scales::percent(params$gtarg)`

# Build Habitat Protection Goals

```{r if-goals, echo = FALSE}
# get reference line
z = params$structure
Mtar = params$gtarg
Atar = 1 - Mtar^(1/z)
A = c(seq(0, 1, by = 0.005))
M = (1-A)^z
preddf = data.frame(A = A, M = M)

# get target goal A=1−M^(1/z)
forplot = data.frame(M = Mtar, A = Atar)

ggplot(data = forplot, aes(x = A, y = M)) +
    geom_line(data = preddf, color = 'lightgreen') +
    geom_point(size = 3, color = 'darkblue') +
    geom_segment(aes(x = Atar, xend = Atar, y = 0, yend = Mtar),
                 linetype = "dashed", color = "lightgray") +
    geom_segment(aes(x = 0, xend = Atar, y = Mtar, yend = Mtar),
                 linetype = "dashed", color = "lightgray") +
    annotate("text", x = Atar, y = 0, label = scales::percent(Atar),
             vjust = 1.5, size = 3.5, color = "black") +
    annotate("text", x = 0, y = Mtar, label = scales::percent(Mtar),
             hjust = -0.2, size = 3.5, color = "black") +
    scale_x_continuous(labels = scales::percent) +
    scale_y_continuous(labels = scales::percent) +
    labs(x = paste0("% of Area lost"),
         y = paste0("% of Segregating sites remained"))
```


# Interpretation

To preserve at least **`r scales::percent(Mtar)`** of genetic diversity intact,
at most **`r scales::percent(Atar)`** of habitat can be lost,
at least **`r scales::percent(1-Atar)`** of habitat needs to be protected,
given a MAR scaling factor z of **`r z`**.

Genetic diversity is measured as allelic richness in this report.
Higher population structure (higher **z**) results in faster genetic diversity loss with habitat loss.
To retain more genetic diversity, more area must be protected.

